---
title: "Parameter Sweep Evaluation"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## Setup

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

# Load libraries
library(RColorBrewer)
library(dittoSeq)
```

```{r specify_paths}
# Set general input paths to all analysis files
analysis.path <- snakemake@params[["out_path"]]

# Set path to masks
masks.path <- snakemake@params[["masks_path"]]
```

```{r helper_fnc}
## General helper fnc.

# Function to read in any of the outputs of CISI training
# file: path to file to read in 
# type: res (results.txt files with correlations), x (X_*.h5ad files with anndata
# object), a (version_*.txt with A/Phi experiment design matrix) or u (gene_modules.csv
# containing the learned dictionary)
read_results <- function(file, type){
  # Read out parameter types of parameter sweep and the values used to create
  # particular output 'file'
  params <- gsub("/[^/]+$", "", gsub(".*parameter_sweep/", "", file))
  params <- str_split(str_split(params, "/")[[1]], "_")
  # Define what type of output of CISI is read in
  if (type=="res"){
    # If type="res" file containing correlation results is read in
    # Column gets added to show if results comes from noisy or no noise simulations
    res <- read_tsv(file, show_col_types=FALSE) %>%
      mutate(simulation=ifelse((grepl("composite", file) | grepl("no_noise", file)), 
                               "no_noise", "noisy"),
             folder=gsub("/[^/]+$", "", file))
  } else if (type=="x") {
    # If type="x", then anndata object gets read into SCE containing either
    # ground truth or simulated/decomposed expressions (normalized and subsetted
    # according to what was specified/used in the CISI run)
    res <- readH5AD(file)
    metadata(res) <- list(ground_truth=ifelse(grepl("X_test", file), 
                                              "ground_truth", 
                                              "simulated"))
  } else if (type=="a"){
    # If type="a" file containing experiment matrix is read in specifying
    # which proteins are measured in the same channel
    res <- read.table(file, row.names=1, sep="\t") %>%
      dplyr::rename_all(~ (stringr::str_replace_all( ., "\\.", "-" ))) %>%
      rownames_to_column(var="channel")
  } else if (type=="u"){
    # If type="u" file containing dictionary is read in
    res <- read.csv(file) %>%
      dplyr::rename(protein=X) %>%
      dplyr::rename_with(~gsub("X", "", .x, fixed=TRUE)) %>%
      melt(id="protein") %>%
      dplyr::rename(module=variable, membership=value)
  }
  
  # Add information about which type of parameters were evaluated in the parameter
  # sweep and what the values were for particular file
  for (p in params){
    if (type=="x"){
      metadata(res)[[p[1]]] <- p[2]
    } else {
      res <- res %>%
        mutate(!!as.symbol(p[1]):=p[2])
    }
  }
  
  res
}
```

## Read Inputs

First, we read in the results from CISI pilot run for the chosen immune markers
(CD15, CD20, CD3, CD38, CD4, CD68, CD8a, ICOS, Ki-67, MPO, panCK, SMA, CD303, 
FOXP3, GranzymeB). The pilot ran was done using no normalization and with k=1.

```{r read_inputs}
## Read results
# Read in all results (correlations) into one dataframe
results.files <- list.files(analysis.path, 'simulation_results.txt',
                            full.names=TRUE, recursive=TRUE)

results.df <- lapply(results.files, read_results, type="res") %>% 
  bind_rows() %>%
  arrange(desc("Gene average"))


## Read X_test and X_simulated
# Specify X_test and X_simulated files from best run 
X.files <- list.files(results.df$folder[1], "X_", full.names=TRUE, recursive=TRUE)

# Read in sce experiments from saved anndata 
X.list <- lapply(X.files, read_results, type="x")
X.list <- lapply(X.list, function(sce.temp){
  assay(sce.temp, "exprs") <- asinh(counts(sce.temp)/1)
  
  sce.temp
})



## Read U
# Specify U file from best run 
u.files <- list.files(results.df$folder[1], "gene_modules.csv",
                      full.names=TRUE, recursive=TRUE)
u <- lapply(u.files, read_results, type="u")


## Read A
# Specify A/Phi file from best run 
a.files <- list.files(results.df$folder[1], "version_*",
                      full.names=TRUE, recursive=TRUE)
a <- lapply(a.files, read_results, type="a") 


# Read in masks for tonsil data
masks <- loadImages(masks.path, as.is=TRUE)
mcols(masks) <- DataFrame(sample_id=names(masks))
```

## Results

### Barplot of Average Gene Correlation

For each different measurement used to test training results, a barplot is shown 
comparing CISI results.

```{r plot_average_gene_correlation, results="asis", fig.height=8, fig.width=18, fig.align="center"}
# For each different measurement of training results, plot barplot 

# Melt dataframe for plotting
data_temp <- results.df %>%
  dplyr::select(-c("dataset", "training", "datasize", "version", "Best crossvalidation fold")) %>%
  pivot_longer(!c("simulation", "normalisation"), names_to="measure", values_to="value")

# Create barplot
results.barplot <- plot_cisi_results(data_temp, "measure", "value", "normalisation")
print(results.barplot)
```

### Per protein results

#### Correlations per Protein *(k=1, arcsinh)* {.tabset}

Correlation between ground truth and decomposed results per protein for  
asinh transformed counts.

```{r results_per_protein_cor, fig.height=6, fig.width=5, fig.align="center", results="asis"}
aoi <- "exprs"
# Calculate correlations between ground truth and simulated data for each protein
X.cor <- lapply(X.list, function(sce){
  counts.long <- as.data.frame(assays(sce)[[aoi]]) %>%
    mutate(protein=rownames(.)) %>%
    melt() %>%
    dplyr::rename(!!as.symbol(metadata(sce)$ground_truth):=value,
                  cell=variable) %>%
    mutate(normalisation=metadata(sce)$normalisation)
  
}) %>% bind_rows() %>%
  group_by(protein, cell, normalisation) %>%
  summarise_all(na.omit) %>%
  group_by(normalisation, protein) %>%
  mutate(correlation=cor(ground_truth, simulated))

# Plot correlations for normalized and combi CISI run
for (i in unique(X.cor$normalisation)) {
  cat("#####", i, "\n")
  
  proteinCorr <- plot_protein_cor(X.cor %>% 
                                    dplyr::filter(normalisation==i)) + ylim(0, 1)
  
  print(proteinCorr)
  cat("\n\n")
}
```

## Image results

#### FOXP3 of Test Image *(k=1, arcsinh)* {.tabset}

Expression values for cells in test image for worst performing protein FOXP3 for
unnormalized and combi CISI run.

```{r plot_FOXP3, fig.height=10, fig.width=10, fig.align="center", fig.keep=c(13, 26), results="asis"}
# Call plot_cells to get individual plots for test roi for decomposed and true
# results
poi <- "FOXP3"
for (n in unique(X.cor$normalisation)){
  cat("#####", n, "\n")
  
  img.foxp3 <- plot_cells(keep(X.list, function(sce.temp){
    metadata(sce.temp)$normalisation == n
  }), masks, poi)
  # Plot decomposed vs true results for test roi (002)
  img.foxp3 <- plot_grid(plotlist=append(img.foxp3[grepl("20220520_TsH_th152_cisi1_002",
                                                         names(img.foxp3))],
                                         img.foxp3[grepl("legend",
                                                         names(img.foxp3))]), ncol=2, 
                         labels=unlist(lapply(X.list, function(sce){metadata(sce)$ground_truth})),
                         label_size=15, hjust=c(-2, -1.5), 
                         vjust=1, scale=0.9)
  print(img.foxp3)
  cat("\n\n")
}
```

#### CD303 of Test Image *(k=1, arcsinh)* {.tabset}

Expression values for cells in test image for second worst performing protein CD303
for unnormalized and combi CISI run.

```{r plot_CD303, fig.height=10, fig.width=10, fig.align="center", fig.keep=c(13, 26), results="asis"}
# Call plot_cells to get individual plots for test roi for decomposed and true
# results
poi <- "CD303"
for (n in unique(X.cor$normalisation)){
  cat("#####", n, "\n")
  
  # Call plot_cells to get individual plots for test roi for decomposed and true
  # results
  img.cd303 <- plot_cells(keep(X.list, function(sce.temp){
    metadata(sce.temp)$normalisation == n
  }), masks, poi)
  # Plot decomposed vs true results for test roi (002)
  img.cd303 <- plot_grid(plotlist=append(img.cd303[grepl("20220520_TsH_th152_cisi1_002",
                                                         names(img.cd303))],
                                         img.cd303[grepl("legend",
                                                         names(img.cd303))]), ncol=2, 
                         labels=unlist(lapply(X.list, function(sce){metadata(sce)$ground_truth})),
                         label_size=15, hjust=c(-2, -1.5), 
                         vjust=1, scale=0.9)
  print(img.cd303)
  cat("\n\n")
}
```

#### CD68 of Test Image *(k=1, arcsinh)* {.tabset}

Expression values for cells in test image for another bad performing protein CD68
for unnormalized and combi CISI run.

```{r plot_CD68, fig.height=10, fig.width=10, fig.align="center", fig.keep=c(13, 26), results="asis"}
# Call plot_cells to get individual plots for test roi for decomposed and true
# results
poi <- "CD68"
for (n in unique(X.cor$normalisation)){
  cat("#####", n, "\n")
  
  # Call plot_cells to get individual plots for test roi for decomposed and true
  # results
  img.cd68 <- plot_cells(keep(X.list, function(sce.temp){
    metadata(sce.temp)$normalisation == n
  }), masks, poi)
  # Plot decomposed vs true results for test roi (002)
  img.cd68 <- plot_grid(plotlist=append(img.cd68[grepl("20220520_TsH_th152_cisi1_002",
                                                       names(img.cd68))],
                                        img.cd68[grepl("legend",
                                                       names(img.cd68))]), ncol=2, 
                        labels=unlist(lapply(X.list, function(sce){metadata(sce)$ground_truth})),
                        label_size=15, hjust=c(-2, -1.5), 
                        vjust=1, scale=0.9)
  print(img.cd68)
  cat("\n\n")
}
```

#### CD38 of Test Image *(k=1, arcsinh)* {.tabset}

Expression values for cells in test image for another bad performing protein CD38
for unnormalized and combi CISI run.

```{r plot_CD38, fig.height=10, fig.width=10, fig.align="center", fig.keep=c(13, 26), results="asis"}
# Call plot_cells to get individual plots for test roi for decomposed and true
# results
poi <- "CD38"
for (n in unique(X.cor$normalisation)){
  cat("#####", n, "\n")
  
  # Call plot_cells to get individual plots for test roi for decomposed and true
  # results
  img.cd38 <- plot_cells(keep(X.list, function(sce.temp){
    metadata(sce.temp)$normalisation == n
  }), masks, poi)
  # Plot decomposed vs true results for test roi (002)
  img.cd38 <- plot_grid(plotlist=append(img.cd38[grepl("20220520_TsH_th152_cisi1_002",
                                                       names(img.cd38))],
                                        img.cd38[grepl("legend",
                                                       names(img.cd38))]), ncol=2, 
                        labels=unlist(lapply(X.list, function(sce){metadata(sce)$ground_truth})),
                        label_size=15, hjust=c(-2, -1.5), 
                        vjust=1, scale=0.9)
  print(img.cd38)
  cat("\n\n")
}
```

#### GranzymeB of Test Image *(k=1, arcsinh)* {.tabset}

Expression values for cells in test image for another bad performing protein GranzymeB
for unnormalized and combi CISI run.

```{r plot_GranzymeB, fig.height=10, fig.width=10, fig.align="center", fig.keep=c(13, 26), results="asis"}
# Call plot_cells to get individual plots for test roi for decomposed and true
# results
poi <- "GranzymeB"
for (n in unique(X.cor$normalisation)){
  cat("#####", n, "\n")
  
  # Call plot_cells to get individual plots for test roi for decomposed and true
  # results
  img.granzymeB <- plot_cells(keep(X.list, function(sce.temp){
    metadata(sce.temp)$normalisation == n
  }), masks, poi)
  # Plot decomposed vs true results for test roi (002)
  img.granzymeB <- plot_grid(plotlist=append(img.granzymeB[grepl("20220520_TsH_th152_cisi1_002",
                                                                 names(img.granzymeB))],
                                             img.granzymeB[grepl("legend",
                                                                 names(img.granzymeB))]), ncol=2, 
                             labels=unlist(lapply(X.list, function(sce){metadata(sce)$ground_truth})),
                             label_size=15, hjust=c(-2, -1.5), 
                             vjust=1, scale=0.9)
  print(img.granzymeB)
  cat("\n\n")
}
```

#### CD20 of Test Image *(k=1, arcsinh)* {.tabset}

Expression values for cells in test image for best performing protein CD20
for unnormalized and combi CISI run.

```{r plot_CD20, fig.height=10, fig.width=10, fig.align="center", fig.keep=c(13, 26), results="asis"}
# Call plot_cells to get individual plots for test roi for decomposed and true
# results
poi <- "CD20"
for (n in unique(X.cor$normalisation)){
  cat("#####", n, "\n")
  
  # Call plot_cells to get individual plots for test roi for decomposed and true
  # results
  img.cd20 <- plot_cells(keep(X.list, function(sce.temp){
    metadata(sce.temp)$normalisation == n
  }), masks, poi)
  # Plot decomposed vs true results for test roi (002)
  img.cd20 <- plot_grid(plotlist=append(img.cd20[grepl("20220520_TsH_th152_cisi1_002",
                                                       names(img.cd20))],
                                        img.cd20[grepl("legend",
                                                       names(img.cd20))]), ncol=2, 
                        labels=unlist(lapply(X.list, function(sce){metadata(sce)$ground_truth})),
                        label_size=15, hjust=c(-2, -1.5), 
                        vjust=1, scale=0.9)
  print(img.cd20)
  cat("\n\n")
}
```

#### CD3 of Test Image *(k=1, arcsinh)* {.tabset}
 
Expression values for cells in test image for CD3, a protein performing mediocre
for unnormalized and combi CISI run.

```{r plot_CD3, fig.height=10, fig.width=10, fig.align="center", fig.keep=c(13, 26), results="asis"}
# Call plot_cells to get individual plots for test roi for decomposed and true
# results
poi <- "CD3"
for (n in unique(X.cor$normalisation)){
  cat("#####", n, "\n")
  
  # Call plot_cells to get individual plots for test roi for decomposed and true
  # results
  img.cd3 <- plot_cells(keep(X.list, function(sce.temp){
    metadata(sce.temp)$normalisation == n
  }), masks, poi)
  # Plot decomposed vs true results for test roi (002)
  img.cd3 <- plot_grid(plotlist=append(img.cd3[grepl("20220520_TsH_th152_cisi1_002",
                                                     names(img.cd3))],
                                       img.cd3[grepl("legend",
                                                     names(img.cd3))]), ncol=2, 
                       labels=unlist(lapply(X.list, function(sce){metadata(sce)$ground_truth})),
                       label_size=15, hjust=c(-2, -1.5), 
                       vjust=1, scale=0.9)
  print(img.cd3)
  cat("\n\n")
}
```

#### CD303 protein distribution *(k=1, arcsinh)* 

```{r CD303_distributions, fig.height=4, fig.width=10, fig.align="center"}
# Have a look at dittoRidgePlots of a bad performing protein
X.list <- lapply(X.list, function(sce.temp) {
  colData(sce.temp)$dummy <- "1"
  
  sce.temp
})

plot.list <- NULL
for(sce in X.list){
  plot.list <- append(plot.list, list(dittoRidgePlot(sce, var="CD303", 
                                                     group.by="dummy", assay="exprs") +
                                        ggtitle(metadata(sce)$ground_truth) +
                                        coord_cartesian(xlim=c(0, 6.2), expand=TRUE)))
}
print(plot_grid(plotlist=plot.list, ncol=1))
```

## Matrix designs 

### U {.tabset}

```{r plot_U, fig.height=8, fig.width=18, fig.align="center", results="asis"}
# Plot U for normalized and combi CISI run
cor <- plot_U(u, "normalisation", "dataset")
```

### A {.tabset}

```{r plot_A, fig.height=8, fig.width=18, fig.align="center", results="asis"}
plot_A(a %>%
         dplyr::select(-training), "normalisation", "dataset")
```