---
title: "Analysis of U stability"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
    pandoc_args: [ 
      "--output=analysis_U_stability_Tonsil_th152.html" 
    ]
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## Setup

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(cowplot)
library(ggplot2)
library(ggpattern)
library(ggsci)
library(utils)
library(reshape2)
library(stringr)
library(gridGraphics)
library(zellkonverter)
library(SingleCellExperiment)
library(cytomapper)
```

```{r specify_paths}
# Set general input paths to all analysis files
analysis.path <- "/mnt/bb_dqbm_volume/analysis/Tonsil_th152"
# analysis.path <- "/mnt/bb_dqbm_volume/analysis/Immucan_lung"

u_stability.path <- file.path(analysis.path, "tests/test_U_stability")
u_stability_par.path <- file.path(analysis.path, "tests/test_U_stability_par")
```

```{r helper_fnc}
# Import helper fncs
source("../analysis/helpers/helper_fnc.R")
```

## Training Data Size

Read in results from the stability analysis of U using different training data sizes 
(U_stability.ipynb). We then look at the correlations between each U and the 
'ground truth', eg. the U computed from the largest training test using the 
[Mantel test] (https://tiagoolivoto.github.io/metan/reference/mantel_test.html).

```{r read_u_stability_size}
# Read .csv file
u_stability.list <- lapply(list.files(u_stability.path, "results.csv", full.names=TRUE, 
                                    recursive=TRUE), function(file){
  k_name <- gsub("k_", "", str_split(file, "/")[[1]][8])
  read_csv(file)[-1] %>% mutate(k=k_name)
})

# Read in segmentation SCE results
u_stability_modules.files <- list.files(u_stability.path, "gene_modules.csv",
                                        full.names=TRUE, recursive=TRUE)
u_stability.modules <- lapply(u_stability_modules.files, read_U, type="size") %>% bind_rows()
```

```{r plot_u_stability_size,  fig.height = 8, fig.width = 18, fig.align = "center", include=FALSE}
# Reshape correlation distances for plotting
u_stability.long <- lapply(u_stability.list, function(stab){
  stab %>%
  melt(id.vars="k") %>%
  dplyr::rename(train_size=variable, distance=value) %>%
  mutate(train_size=strtoi(train_size))
}) %>% bind_rows()
  
# Plot stability of U
ggplot(u_stability.long, aes(train_size, distance)) +
  geom_point(aes(color=k, fill=k)) +
  geom_smooth(aes(color=k, fill=k), method="lm") +
  theme_cowplot(title.fontsize) +
  scale_color_npg() +
  labs(x="Training size", y="Correlation distance")
```

### Mantel Test for k=3 and k=1 *(d=80, iter=10, paper_norm)*

```{r plot_u_stability_size_mantel, fig.height = 8, fig.width = 18, fig.align = "center"}
# Calculate mantel test for each U compared to 'ground truth' U of average U
# using the full data set for all tested k
u_stability.mantel <- lapply(unique(u_stability.modules$k), function(k_iter){
  # Empty list to save correlation matrices
  u_stability.cor <- list()
  
  # Go over every training size and replicate
  for(s in unique(u_stability.modules$size)){
    temp_list <-  list()
    for(r in unique(u_stability.modules$rep)) {
      res.temp <- u_stability.modules %>%
        dplyr::filter(rep==r & k==k_iter & size==s) %>%
        dplyr::select(-c("rep", "k", "size")) %>%
        pivot_wider(names_from=module, values_from=membership) %>%
        column_to_rownames(var="protein") %>%
        as.matrix()
      
      # Calculate correlation matrix
      temp_list[[length(temp_list)+1]] <- cor(t(res.temp))
    }
    
    u_stability.cor[[length(u_stability.cor)+1]] <- temp_list
  }
  names(u_stability.cor) <- unique(u_stability.modules$size)
  
  # Ground truth
  ground_truth.size <- as.character(max(as.integer(names(u_stability.cor))))
  ground_truth <- Reduce('+', u_stability.cor[[ground_truth.size]]) /
    length(u_stability.cor[[ground_truth.size]])
  
  # Compute mantel test for every U
  mantel_temp <- rapply(u_stability.cor, 
                        function(x, gt) {mantel_test(x, gt)$mantel_r}, 
                        how="list", gt=ground_truth)
  mantel_temp <- as.data.frame(do.call(cbind, mantel_temp)) %>%
    mutate(k=k_iter) %>%
    mutate_all(unlist)

  mantel_temp
}) %>% bind_rows() %>%
  melt(id.vars="k") %>%
  dplyr::rename(size=variable,
                mantel_cor=value) 

u_stability.mantel <- u_stability.mantel %>%
  mutate(size=strtoi(u_stability.mantel$size),
         mantel_cor=1-mantel_cor)

# Plot stability of U using mantel test results
u_stability.mantelPlot <- ggplot(u_stability.mantel, aes(size, mantel_cor)) +
  geom_point(aes(color=k, fill=k)) +
  geom_smooth(aes(color=k, fill=k), method="lm", formula=y ~ poly(x, 2)) +
  theme_cowplot(title.fontsize) +
  scale_color_npg() +
  labs(x="Training size", y=expression(1 - cor["Mantel"]))
u_stability.mantelPlot
```

## SMAF Parameters

Read in results from the stability analysis of U computed using different parameters
(U_stability.ipynb). We then look at the heatmap of U and the pairwise correlations 
between U's using the Mantel test.

### k (W sparsity) *(d=80, iter=10, paper_norm)*

```{r read_u_stability_k}
# Define path for results of different k's (sparsity)
u_stability_k.path <- file.path(u_stability_par.path, "k")

# Read in segmentation SCE results
u_stability_modules_k.files <- list.files(u_stability_k.path, 'gene_modules.csv',
                                         full.names=TRUE, recursive=TRUE)
u_stability_k.modules <- lapply(u_stability_modules_k.files, read_U, type="par") %>% bind_rows()
```

#### Heatmap of U for different k's {.tabset}

```{r plot_u_stability_k, results="asis", fig.height = 8, fig.width = 18, fig.align = "center"}
# Plot U for different k's
u_stability_k.cor <- plot_U(u_stability_k.modules, "k", "rep")
```

```{r plot_u_stability_k_dist,  fig.height = 8, fig.width = 18, fig.align = "center", include=FALSE}
u_stability_k.boxplot <- plot_dist_boxplot(u_stability_k.cor, "k")
u_stability_k.boxplot
```

#### Mantel test for different k's {.tabset}

```{r plot_u_stability_k_mantel, results="asis", fig.height = 8, fig.width = 18, fig.align = "center"}
plot_metan_mantel(u_stability_k.cor, "k")
```

#### {-}

The boxplot shows a summary of the above mantel test results, by only looking at the pairwise correlations between any two matrices for each k.

```{r plot_u_stability_k_mantel_boxplot, fig.height = 8, fig.width = 18, fig.align = "center"}
u_stability_k.mantelBoxplot <- plot_mantel_boxplot(u_stability_k.cor, "k")
print(u_stability_k.mantelBoxplot)
```

### maxItr (number of iterations to compute U) *(k=3, d=80, paper_norm)*

```{r read_u_stability_maxItr}
# Define path for results of different max iterations when computing dictionary U
u_stability_maxItr.path <- file.path(u_stability_par.path, "maxItr")

# Read in segmentation SCE results
u_stability_modules_maxItr.files <- list.files(u_stability_maxItr.path, 'gene_modules.csv',
                                               full.names=TRUE, recursive=TRUE)
u_stability_modules_maxItr.modules <- lapply(u_stability_modules_maxItr.files, read_U, type="par") %>% 
  bind_rows() %>%
  dplyr::rename(maxItr=k)
```

#### Heatmap of U for different maxItr {.tabset}

```{r plot_u_stability_maxItr, results="asis", fig.height = 8, fig.width = 18, fig.align = "center"}
# Plot U for different maxItr
u_stability_maxItr.cor <- plot_U(u_stability_modules_maxItr.modules, "maxItr", "rep")
```

```{r plot_u_stability_maxItr_dist, fig.height = 8, fig.width = 18, fig.align = "center", include=FALSE}
u_stability_maxItr.boxplot <- plot_dist_boxplot(u_stability_maxItr.cor, "maxItr")
u_stability_maxItr.boxplot
```

#### Mantel test for different maxItr {.tabset}

```{r plot_u_stability_maxItr_mantel, results="asis", fig.height = 8, fig.width = 18, fig.align = "center"}
plot_metan_mantel(u_stability_maxItr.cor, "maxItr")
```

#### {-}

The boxplot shows a summary of the above mantel test results, by only looking at the pairwise correlations between any two matrices for each maxItr.

```{r plot_u_stability_maxItr_mantel_boxplot, fig.height = 8, fig.width = 18, fig.align = "center"}
u_stability_maxItr.mantelBoxplot <- plot_mantel_boxplot(u_stability_maxItr.cor, "maxItr")
print(u_stability_maxItr.mantelBoxplot)
```

### Transformation of X (before SMAF training) *(k=1, d=80, iter=10)*

```{r read_u_stability_xform}
# Define path for results of different transformations before training
u_stability_xform.path <- file.path(u_stability_par.path, "transformation")

# Read in segmentation SCE results
u_stability_modules_xform.files <- list.files(u_stability_xform.path, 'gene_modules.csv',
                                              full.names=TRUE, recursive=TRUE)
u_stability_xform.modules <- lapply(u_stability_modules_xform.files, read_U, type="par") %>% 
  bind_rows() %>%
  dplyr::rename(transformation=k)
```

#### Heatmap of U for different transformations {.tabset}

```{r plot_u_stability_xform, results="asis", fig.height = 8, fig.width = 18, fig.align = "center"}
# Plot U for different transformations
u_stability_xform.cor <- plot_U(u_stability_xform.modules, "transformation", "rep")
```

```{r plot_u_stability_xform_dist,  fig.height = 8, fig.width = 18, fig.align = "center", include=FALSE}
u_stability_xform.boxplot <- plot_dist_boxplot(u_stability_xform.cor, "transformation", 
                                               to_int=FALSE)
u_stability_xform.boxplot
```

#### Mantel test for different transformations {.tabset}

```{r plot_u_stability_xform_mantel, results="asis", fig.height = 8, fig.width = 18, fig.align = "center"}
plot_metan_mantel(u_stability_xform.cor, "transformation")
```

#### {-}

The boxplot shows a summary of the above mantel test results, by only looking at the pairwise correlations between any two matrices for each transformation.

```{r plot_u_stability_xform_mantel_boxplot, fig.height = 8, fig.width = 18, fig.align = "center"}
u_stability_xform.mantelBoxplot <- plot_mantel_boxplot(u_stability_xform.cor, "transformation", 
                                                       to_int=FALSE)
print(u_stability_xform.mantelBoxplot)
```

### Dictionary size of U (number of modules in U) *(k=1, iter=10, paper_norm)*

```{r read_u_stability_dict}
# Define path for results of different dictionary sizes
u_stability_dict.path <- file.path(u_stability_par.path, "dictSize")

# Read in segmentation SCE results
u_stability_modules_dict.files <- list.files(u_stability_dict.path, 'gene_modules.csv',
                                              full.names=TRUE, recursive=TRUE)
u_stability_dict.modules <- lapply(u_stability_modules_dict.files, read_U, type="par") %>% 
  bind_rows() %>%
  dplyr::rename(dictionary_size=k)
```

#### Heatmap of U for different dictionary sizes {.tabset}

```{r plot_u_stability_dict, results="asis", fig.height = 8, fig.width = 18, fig.align = "center"}
# Plot U for different k's
u_stability_dict.cor <- plot_U(u_stability_dict.modules, "dictionary_size", "rep")
```

```{r plot_u_stability_dict_dist,  fig.height = 8, fig.width = 18, fig.align = "center", include=FALSE}
u_stability_dict.boxplot <- plot_dist_boxplot(u_stability_dict.cor, "dictionary_size", 
                                               to_int=FALSE)
u_stability_dict.boxplot
```

#### Mantel test for different dictionary sizes {.tabset}

```{r plot_u_stability_dict_mantel, results="asis", fig.height = 8, fig.width = 18, fig.align = "center"}
plot_metan_mantel(u_stability_dict.cor, "dictionary_size")
```

#### {-}

The boxplot shows a summary of the above mantel test results, by only looking at the pairwise correlations between any two matrices for each dictionary size.

```{r plot_u_stability_dict_mantel_boxplot, fig.height = 8, fig.width = 18, fig.align = "center"}
u_stability_dict.mantelBoxplot <- plot_mantel_boxplot(u_stability_dict.cor, "dictionary_size", 
                                                       to_int=TRUE)
print(u_stability_dict.mantelBoxplot)
```
