---
title: "Proof-of-Concept Experiment Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


## Setup
***

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

# Load libraries
```

```{r helper_fnc}
# Import helper fncs
source("../analysis/helpers/helper_fnc.R")
```

```{r specify_paths}
## TODO: Change paths
# Set general input paths to all analysis files
analysis.path <- "/mnt/bb_dqbm_volume/analysis/...."

# Set path to masks
masks.path <- "/mnt/bb_dqbm_volume/data/..."
```

### Plot defaults

Next, we set some defaults for consisistend plotting accross the script.

```{r set_plot_defaults}
## TODO: set celltypes and colours
# Define celltype colours used by the rest of the script
celltype.col <- c(brewer.pal(12, "Paired"),
                           brewer.pal(8, "Pastel2")[-c(3,5,8)],
                           brewer.pal(12, "Set3")[-c(2,3,8,9,11,12)])
names(celltype.col) <- c()
```


## Read Inputs
***

First, we read in the results from the proof-of-concept experiment using the 
SABER-technology to create composite measurements from a pre-defined A, 
decomposed using the best parameters from the parameter sweep and its predicted 
dictionary.

```{r read_inputs}
## Read results
# Read in all results into one dataframe
results.files <- list.files(analysis.path, 'simulation_results.txt',
                            full.names=TRUE, recursive=TRUE)
results.df <- lapply(results.files, read_single_correlation_results) %>% 
  bind_rows() 


## Read X_test and X_simulated
# Specify X_test and X_simulated files to be read in
X.files <- list.files(analysis.path, "X_", full.names=TRUE, recursive=TRUE)

# Read in all sce experiments from saved anndata and save additional info in metadata
# (e.g. if it is the ground truth or decomposed X)
X.list <- lapply(X.files, read_single_anndata)
X.list <- lapply(X.list, function(sce.temp){
  assay(sce.temp, "exprs") <- asinh(counts(sce.temp)/1)
  
  sce.temp
})
names(X.list) <- unlist(lapply(X.list, function(sce.temp){metadata(sce.temp)$ground_truth}))


## Read U
u.files <- list.files(analysis.path, "gene_modules.csv",
                      full.names=TRUE, recursive=TRUE)
u <- lapply(u.files, read_single_U)


## Read A
a.files <- list.files(analysis.path, "version_*",
                      full.names=TRUE, recursive=TRUE)
a <- lapply(a.files, read_single_A) 


# Read in masks 
masks <- loadImages(masks.path, as.is=TRUE)
mcols(masks) <- DataFrame(sample_id=names(masks))
```


## Results
***

### Barplot of Results 

The barplot shows the different measurements used to evaluate decomposition of CISI
with the specified U, A and parameters.

```{r plot_results, fig.asp=0.4, fig.width=10, fig.align="center"}
# For each different measurement of training results, plot barplot 

# Melt dataframe for plotting
data_temp <- results.df %>%
  dplyr::select(-c("dataset", "training", "datasize", "version", "Best crossvalidation fold")) %>%
  pivot_longer(!c("simulation"), names_to="measure", values_to="value")

# Create barplot
results.barplot <- plot_cisi_results(data_temp, "measure", "value", "simulation")
print(results.barplot)
```

### Per Protein Results
***

#### Correlations per Protein *(arcsinh)*

Correlation between ground truth and decomposed results per protein for  
asinh transformed counts.

```{r results_per_protein_cor, fig.asp=1, fig.width=8, fig.align="center"}
aoi <- "exprs"
# Calculate correlations between ground truth and simulated data for each protein
X.cor <- lapply(X.list, function(sce){
  counts.long <- as.data.frame(assays(sce)[[aoi]]) %>%
    mutate(protein=rownames(.)) %>%
    melt() %>%
    dplyr::rename(!!as.symbol(metadata(sce)$ground_truth):=value,
                  cell=variable)
}) %>% bind_rows() %>%
  group_by(protein, cell) %>%
  summarise_all(na.omit) %>%
  group_by(protein) %>%
  mutate(correlation=cor(ground_truth, simulated)) %>%
  ungroup()

# Plot correlations per protein
proteinCorr <- plot_protein_cor(X.cor %>%
                                  dplyr::select(protein, correlation) %>%
                                  distinct()) + ylim(0, 1)
```

#### Scatterplots of Counts *(arcsinh)*

Scatterplot of ground truth vs decomposed results per protein for arcsinh 
transformed counts.

```{r results_per_protein_scatter, fig.asp=1.2, fig.width=10, fig.align="center"}
# Plot for scatter plot of true vs decomposed, add diagonal and regression line, 
# as well as R (pearson correlation coefficient)
proteinPlot <- plot_protein_scatterplot(X.cor)
  
print(proteinPlot)
```

#### Scatterplots of a Combination of Two Proteins *(arcsinh)* {.tabset}

In the next part, we look at the combination of arcsinh transformed counts for two 
proteins. This is especially important for manual gating for celltypes. 

```{r results_per_protein_combined_scatter, fig.asp=0.5, fig.width=10, fig.align="center", results="asis"}
pcoi <- list(list("CD3", "CD20"), list("CD3", "CD4"), list("CD3", "CD8"))

for (combi in pcoi) {
  cat("##### ", combi[1], "vs", combi[2], "\n")

  # Plot counts of proteins of interest of decomposed and true 
  # datasets coloured by different celltypes
  proteinPlotCombi <- plot_exprs(X.cor, "", combi[1], combi[2])
  print(proteinPlotCombi)
  cat("\n\n")
}
```

### Image Results
***

#### Espression Values for All Proteins *(arcsinh)* {.tabset}

Expression values for cells in image for all proteins. 

```{r plot_protein_expr, fig.asp=0.5, fig.width=18, fig.align="center", fig.keep=c(13, 26), results="asis"}
# Get the names of all proteins in the experiment
pois <- unique(X.cor$protein)

# Call plot_cells to get individual plots for test roi for decomposed and true
# results
for (p in pois){
  cat("#####", p, "\n")
  
  img <- plot_cells(X.list, masks, p)
  # Plot decomposed vs true results
  img <- plot_grid(plotlist=append(img[!grepl("legend",
                                             names(img))],
                                   img[grepl("legend",
                                             names(img))]), ncol=2, 
                   labels=unlist(lapply(X.list, function(sce){metadata(sce)$ground_truth})),
                   label_size=15, hjust=c(-2, -1.5), 
                   vjust=1, scale=0.9)
  print(img)
  cat("\n\n")
}
```

### Clustering Results
***

To investigate the clustering results of ground truth and simulated data a bit more, 
we show the overlap of clusters computed using Rphenograph (run with k=100 clusters)
individually and compare which clusters overlap using a heatmap.

```{r clustering}
# Compute clusters using Rphenograph 
all.clusters <- lapply(X.list, function(sce){
  mat <- t(assay(sce, "exprs"))
  out <- Rphenograph(mat)
  sce$clusters <- factor(membership(out[[2]]))
  
  clusters <- as.vector(membership(out[[2]])) %>%
    as.data.frame() %>%
    dplyr::rename(!!as.symbol(metadata(sce)$ground_truth):=".") %>%
    mutate(cell=colData(sce)$cell_id) 

  clusters
}) %>% bind_rows() %>%
  group_by(cell) %>%
  summarise_all(na.omit) %>%
  ungroup()
```

#### Cluster annotation *(arcsinh)*

To identify celltypes according to the clustering, we next have a look at the protein 
expression of the different clusters and manually add the celltypes accordingly. 

```{r plot_clustering_vs_expression, fig.asp=0.5, fig.width=10, fig.align="center"}
ht_list <- NULL
for (n in names(X.list)){
  ht_list <- ht_list +
    dittoHeatmap(X.list[[n]], 
                 #genes=rownames(spe)[rowData(spe)$use_channel],
                 assay="exprs", scale="none",
                 heatmap.colors=magma(100), 
                 annot.by=c("clusters"),
                 annot.colors=celltype.col,
                 complex=TRUE)
}

# Draw heatmaps next to each other
draw(ht_list)

# Add celltypes
celltypes_to_clusters <- c()
```

#### Cluster Overlap between Ground Truth and Decomposed Data *(arcsinh)*

```{r plot_clustering, fig.asp=0.5, fig.width=10, fig.align="center"}
# Plot heatmaps showing overlap of clusters 
cluster.matrix <- table(paste("sim", all.clusters) %>% 
                          pull(simulated), 
                        paste("gt", all.clusters %>% 
                                pull(ground_truth))) %>% 
  as.matrix()
cluster.matrix <- log10(cluster.matrix + 10)


# Create heatmap
cluster.overlap <- Heatmap(cluster.matrix,
                           show_row_dend=FALSE,
                           show_column_dend=FALSE,
                           col=magma(100),
                           column_title="Cluster Overlap",
                           row_names_gp=gpar(fontsize=axis_title.fontsize),
                           column_names_gp=gpar(fontsize=axis_title.fontsize),
                           column_title_gp=gpar(fontsize=title.fontsize),
                           rect_gp=gpar(col="white", lwd=1)) 

print(cluster.overlap)
```


#### ARI *(arcsinh)*

```{r plot_ARI, fig.asp=1, fig.width=7, fig.align="center"}
```


## Matrix designs
***

### U 

```{r plot_U, fig.asp=0.7, fig.width=8, fig.align="center"}
# Plot used U
u.plot <- plot_single_U(u, paste0("U\n(dictionary size: ", length(unique(u$module)), ")"))
print(u.plot)
```

### A 

```{r plot_A, fig.asp=0.6, fig.width=8, fig.align="center"}
# Plot used A
a.plot <- plot_single_A(a, "A")
print(a.plot)
```