---
title: "Analysis of Immucan Lung"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
    keep_md: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## Setup

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

# Load libraries
library(RColorBrewer)
```

```{r specify_paths}
# Set general input paths to all analysis files
analysis.path <- "/mnt/bb_dqbm_volume/analysis/Immucan_lung/training/full"

## Input data
data.path <- "/mnt/bb_dqbm_volume/data/Immucan_lung"

# Set path to masks
masks.path <- file.path(data.path, "masks")
# Set path to random forest classifier
rf.path <- file.path(data.path, "rffit_v7_all_markers.rds")
```

```{r helper_fnc}
# Import helper fncs
source("../analysis/helpers/helper_fnc.R")
```

## Read Inputs

```{r read_inputs}
## Read results
# Read in all results for tonsil and lung comparison into one dataframe
results.files <- list.files(analysis.path, 'simulation_results.txt',
                            full.names=TRUE, recursive=TRUE)
results.files <- results.files[grepl("normalized", results.files)]

results.df <- lapply(results.files, read_results, type="res", voi="norm") %>% 
  bind_rows() %>%
  mutate(norm=ifelse(grepl("combi", norm), "combi", norm))


## Read X_test and X_simulated
# Specify X_test and X_simulated files to be read in
X.files <- list.files(analysis.path, "X_", full.names=TRUE, recursive=TRUE)
X.files <- X.files[grepl("normalized", X.files)]

# Read in all sce experiments from saved anndata and save additional info in metadata
# (e.g. which dataset is used in training and testing, if spillover correction was used 
# and if it is the ground truth or simulated X)
X.list <- lapply(X.files, read_results, type="x", voi="norm")
X.list <- lapply(X.list, function(sce.temp){
  if (grepl("combi", metadata(sce.temp)$norm)){
    metadata(sce.temp)$norm <- "combi"
  } 
  assay(sce.temp, "exprs") <- asinh(counts(sce.temp)/1)
  
  sce.temp
})


# Read in masks for lung data
masks <- loadImages(masks.path, as.is=TRUE)
mcols(masks) <- DataFrame(sample_id=names(masks))

## Read random forest classifier
# Read RF classifier for the lung dataset
rffit.lung <- readRDS(rf.path)
```

## Barplot of Results *(k=1, argsinh)*

For each different measurement used to test training results, a barplot is shown 
comparing CISI results using the original normalization (across cells in training, 
simulating and to what it is compared to) and not using it at all.

```{r plot_results, results="asis", fig.height=8, fig.width=18, fig.align="center"}
# For each different measurement of training results, plot barplot comparing
# spillover corrected and uncorrected results

# Melt dataframe for plotting
data_temp <- results.df %>%
  dplyr::select(-c("dataset", "training", "datasize", "version", "Best crossvalidation fold")) %>%
  pivot_longer(!c("norm", "simulation"), names_to="measure", values_to="value")

# Create barplot
results.barplot <- plot_cisi_results(data_temp, "measure", "value", "norm")
print(results.barplot)
```

## Scatterplot Results with and without Normalization in Training {.tabset}

The following plots show scatterplots of different proteins of interest where cells
are coloured according to celltypes of interest for Immucan lung dataset once trained
on normalized and once trained on unnormalized datasets.

```{r plot_corrections, fig.height=15, fig.width=10, fig.align="center", results="asis"}
names(X.list) <- lapply(X.list, function(x){metadata(x)$ground_truth})
X.list <- append(X.list[-c(1:2)], X.list[1:2])

# List of combinations of proteins of interest
poi <- list(c("CD20", "CD3", "B"), c("VISTA", "CD3", ""))
for (p_vec in poi) {
  cat("###", p_vec[1], "vs", p_vec[2], "\n")
  
  # Plot asinh transformed counts of proteins of interest of decomposed data trained
  # with and without normalization
  X.spillover <- plot_grid(plot_exprs(X.list[1:2], p_vec[3], p_vec[1], p_vec[2]),
                           plot_exprs(X.list[3:4], p_vec[3], p_vec[1], p_vec[2]),
                           plot_exprs(X.list[5:6], p_vec[3], p_vec[1], p_vec[2]),
                           ncol=1, labels=unique(unlist(lapply(X.list, function(sce){
                             metadata(sce)$norm}))),
                           label_size=15, hjust=c(-4.5, -3, -7.5), vjust=0.9, scale=0.9)
  
  print(X.spillover)
  cat("\n\n")
}
```

## Cell Phenotyping *(k=1, arcsinh)* {.tabset}

Next, we use the gating labels of a pre-trained random forest classifier for the 
Immucan lung dataset to newly train a random forest classifier on our decomposed 
argsinh transformed data and predict all celltypes.
We then compare these results to the 'ground truth' celltypes (celltypes predicted from
ground truth data) by calculating the precision, sensitivity and specififity.

```{r rffit, fig.height=8, fig.width=8, fig.align="center", results="asis"}
## Fix colours for plotting
celltype.col <- c(brewer.pal(12, "Paired"),
                           brewer.pal(8, "Pastel2")[-c(3,5,8)],
                           brewer.pal(12, "Set3")[-c(2,3,8,9,11,12)])
celltype.col <- c(celltype.col[1:length(unique(colData(X.list[[1]])$celltype))],
                  "black")
names(celltype.col) <- c(levels(colData(X.list[[1]])$celltype), NA)


# Predict cell type using pretrained classifier on decomposed results and
# compute confusion matrix using celltypes of true data

# Extract decomposted datasets
lungSim.list <- keep(X.list, function(x){
  metadata(x)$ground_truth=="simulated"})

# Train RF classifier
rffit.sim <- lapply(lungSim.list, 
  train_rf, rffit.original=rffit.lung)

# saveRDS(rffit.sim, file.path(analysis.path, "rffit_sim_data_norm.rds"))
# rffit.sim <- readRDS(file.path(analysis.path, "rffit_sim_data_norm.rds"))

# Calculate confusion matrix
lungSim.cm <- lapply(as.list(1:length(lungSim.list)), function(i){
  compute_confusionMatrix(lungSim.list[[i]], rffit.sim[[i]])})

# Plot prediction results/confusion matrix as shown in IMCDataAnalysis script
i <- 1
for (cm in lungSim.cm){
  cat("###", metadata(lungSim.list[[i]])$norm, "\n")
  
  pred.plot <- data.frame(cm$byClass) %>%
    mutate(class=sub("Class: ", "", rownames(cm$byClass))) %>%
    ggplot() +
    geom_point(aes(1 - Specificity, Sensitivity,
                   size=Detection.Rate,
                   fill=class),
               shape=21) +
    scale_fill_manual(values=celltype.col) +
    theme_cowplot(10) +
    ylab("Sensitivity (TPR)") +
    xlab("1 - Specificity (FPR)")
  
  i <- i+1

  print(pred.plot)
  cat("\n\n")
}
```

## Cell classification probabilities *(k=1, arcsinh)* {.tabset}

Next, we have a look at which celltypes the RF predicts for wrongly identified
cells using the classification probabilities per celltype class.

```{r rffit_prob, fig.height=10, fig.width=10, fig.align="center", results="asis"}
# Predict celltype probabilities
lungSim.prob <- lapply(as.list(1:length(lungSim.list)), function(i){
  compute_celltypeProb(lungSim.list[[i]], rffit.sim[[i]])})

# Plot prediction results/confusion matrix as shown in IMCDataAnalysis script
i <- 1
for (p in lungSim.prob){
  cat("###", metadata(lungSim.list[[i]])$norm, "\n")
  
  pred_prob.plot <- p %>%
    pivot_longer(cols=B:Tumor_Ki67) %>%
    ggplot() +
    geom_boxplot(aes(x=name, y=value, fill=name), outlier.size=0.5) +
    facet_wrap(.~truth, ncol=1) + 
    scale_fill_manual(values=celltype.col) +
    theme(panel.background=element_blank(), 
          axis.text.x=element_text(angle=45, hjust=1)) +
    theme_cowplot(10) 
  
  i <- i+1

  print(pred_prob.plot)
  cat("\n\n")
}
```

## Images with celltypes *(k=1, arcsinh)* 

We also have a look at images coloured by celltypes identified using the ground truth
and decomposed/simulated data.

### Image 1 {.tabset}

```{r celltype_images, fig.height=20, fig.width=8, fig.align="center", results="asis", fig.keep=c(5, 10, 15)}
lungSim.list <- lapply(as.list(1:length(lungSim.list)), function(i){
  sce <- lungSim.list[[i]]
  
  cell_class <- data.frame(cell_id=rownames(lungSim.prob[[i]]),
                           celltype_pred=factor(colnames(lungSim.prob[[i]])[max.col(
                             lungSim.prob[[i]] %>%
                               dplyr::select(-truth), ties.method="first")]))
  cell_class[rowMax(as.matrix(lungSim.prob[[i]] %>%
                                dplyr::select(-truth))) < 0.4, "celltype_pred"] <- "uncertain"
  
  # Store labels in SpatialExperiment onject
  colData(sce) <- merge(colData(sce), cell_class, all.x=TRUE)
  colData(sce)$celltype_NA <- colData(sce)$celltype
  colData(sce)$celltype_NA[is.na(colData(sce)$celltype_pred)] <- NA
  
  sce
})

# Specify image index of interest
im <- 1

# Print image im of ground truth celltypes vs decomposed celltypes
for (sce in lungSim.list){
  cat("####", metadata(sce)$norm, "\n")
  
  images <- plot_celltype_images(sce, masks, im, celltype.col)
  
  print(images)
  cat("\n\n")
}
```

### Image 2 {.tabset}

```{r celltype_images2, fig.height=20, fig.width=8, fig.align="center", results="asis", fig.keep=c(5, 10, 15)}
# Specify image index of interest
im <- 10

# Print image im of ground truth celltypes vs decomposed celltypes
for (sce in lungSim.list){
  cat("####", metadata(sce)$norm, "\n")
  
  images.2 <- plot_celltype_images(sce, masks, im, celltype.col)
  
  print(images.2)
  cat("\n\n")
}
```

### Image 3 {.tabset}

```{r celltype_images3, fig.height=20, fig.width=8, fig.align="center", results="asis", fig.keep=c(5, 10, 15)}
# Specify image index of interest
im <- 100

# Print image im of ground truth celltypes vs decomposed celltypes
for (sce in lungSim.list){
  cat("####", metadata(sce)$norm, "\n")
  
  images.3 <- plot_celltype_images(sce, masks, im, celltype.col)
  
  print(images.3)
  cat("\n\n")
}
```

## Cell Phenotyping per ROI *(k=1, arcsinh)* {.tabset}

Next, we compare these results to the 'ground truth' celltypes (celltypes predicted from
ground truth data) by calculating the precision, sensitivity and specififity but this
time per ROI.

```{r rffit_per_ROI, fig.height=8, fig.width=8, fig.align="center", results="asis", include=FALSE, eval=FALSE}
# Plot prediction results/confusion matrix as shown in IMCDataAnalysis script
i <- 1
for (sce in lungSim.list){
  cat("###", metadata(sce)$norm, "\n")
  
  cm.samples <- lapply(unique(colData(lungSim.list[[1]])$sample_id), function(sce){
    cm <- confusionMatrix(data=!is.na(colData(sce)["celltype_pred", "sample_id"==sample]),
                          reference=!is.na(colData(sce)["celltype_NA", "sample_id"==sample]),
                          mode="everything")$byClass %>%
      as.data.frame() %>%
      mutate(class=sub("Class: ", "", rownames(cm$byClass)),
             sample=sample)
    
    cm
  }) %>% bind_rows()
  
  cm

  cat("\n\n")
}
```