---
title: "Proof-of-Concept Analysis using Antibody Mixture"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
    pandoc_args: [ 
      "--output=analysis_tonsil_POC_abmix_2.html" 
    ]
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


## Setup
***

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

# Load libraries
library(batchelor)
```

```{r helper_fnc}
# Import helper fncs
source("../analysis/helpers/helper_fnc.R")
```

```{r specify_paths}
## TODO: Change paths
# Set general input paths to all analysis files
analysis.path <- "/mnt/bb_dqbm_volume/analysis/Tonsil_th179_2"

# Set path to masks
masks.path <- "/mnt/bb_dqbm_volume/data/20221108_TsH_LSS_cisiabmix2_179/steinbock/masks_deepcell"
```


## Read Inputs
***

First, we read in the results from the CISI analysis using a pre-defined A and
the best parameters from the parameter sweep. For this we have data from simulated 
composite data from training, from the actual experiment and the decomposed data.

The channels from the experiment were decomposed using the dictionary U with the 
best predicted performance (from the run using 'min strategy' for A) for a 8 to 6 
compression. This dictionary came from the run with the parameters: k=2, d=10 and 
m=6.

```{r read_inputs}
## Read results
# Read in all results into one dataframe
results.files <- list.files(analysis.path, 'simulation_results.txt',
                            full.names=TRUE, recursive=TRUE)
results.df <- lapply(results.files, read_results, type="res", voi="comparison") %>% 
  bind_rows() %>%
  dplyr::select(-c("dataset", "training", "datasize"))


## Read X_test, X_simulated and X_decomposed 
# Specify files to be read in
X_results.files <- list.files(analysis.path, "X_", full.names=TRUE, recursive=TRUE)

# Read in all sce experiments from saved anndata and save additional info in metadata
# (e.g. if it is the ground truth or simulated X and if it comes from training simulations,
# experiment simulation or the acutal decomposed experiment)
X_results.list <- lapply(X_results.files, read_results, type="x", voi="comparison")
X_results.list <- lapply(X_results.list, function(sce.temp){
  metadata(sce.temp) <- within(metadata(sce.temp), rm(dataset, training)) 
  assay(sce.temp, "exprs") <- asinh(counts(sce.temp)/1)
  
  sce.temp
})


## Read anndata object containing composite measurements
X_simulation.files <- list.files(analysis.path, "composite_measurements.h5ad", 
                             full.names=TRUE, recursive=TRUE)
X_simulation.list <- lapply(X_simulation.files, read_results, type="x", use_voi=FALSE)
# Remove non-composite channels, add if data comes form simulated or real composite 
# measurements and add arcsinh transformed counts
X_simulation.list <- lapply(1:length(X_simulation.list), function(i){
  sce.temp <- X_simulation.list[[i]]
  sce.temp <- sce.temp[grepl("CC\\d_", rownames(sce.temp)), ]
  metadata(sce.temp) <- within(metadata(sce.temp), rm(dataset, training))
  metadata(sce.temp)$ground_truth <- ifelse(grepl("simulated", X_simulation.files[i]),
                                            "simulated", "ground_truth")
  assay(sce.temp, "exprs") <- asinh(counts(sce.temp)/1)
  
  sce.temp
})


## Read U
u.files <- list.files(analysis.path, "gene_modules",
                      full.names=TRUE, recursive=TRUE)
u.files <- u.files[!grepl("min", u.files)]
u <- read_single_U(u.files)


## Read A
a.files <- list.files(analysis.path, "version_*",
                      full.names=TRUE, recursive=TRUE)
a <- read_single_A(a.files) 


# Read in masks 
masks <- loadImages(masks.path, as.is=TRUE)
mcols(masks) <- DataFrame(sample_id=names(masks))
```


## Results
***

### Barplot of Results 

The barplot shows the different measurements used to evaluate decomposition of CISI
with the specified U, A and parameters.

There are three different results: training (decomposed simulated data from training 
compared to ground truth training data), experiment-simulation (decomposed simulated 
data from experiment data compared to ground truth experiment data) and experiment 
(decomposed real composite measurements from experiment compared to ground truth 
experiment data).

```{r plot_results, results="asis", fig.asp=0.5, fig.width=18, fig.align="center"}
# For each different measurement of training results, plot barplot 
# Melt dataframe for plotting
data_temp <- results.df %>%
  dplyr::select(-c("version")) %>%
  pivot_longer(!c("simulation", "comparison"), names_to="measure", values_to="value")

# Create barplot
results.barplot <- plot_cisi_results(data_temp, "measure", "value", "comparison")
print(results.barplot)
```

## Simulated vs Real Composite Measurements
***

### Images {.tabset}

Here we show the composite measurements vs the simulated composite measurements for 
all images. 

```{r plot_images_composite, fig.asp=0.25, fig.width=18, fig.align="center", fig.keep=c(3, 6, 9, 12, 15, 18), results="asis"}
# Find the names of all composite channels
pois <- rownames(X_simulation.list[[1]])
names(pois) <- rownames(X_simulation.list[[1]])


# Call plot_cells to get individual plots for all rois for simulated and real
# composite measurements
for (n in names(pois)){
  cat("####", n, "\n")
  
  img <- plot_cells(X_simulation.list, masks, pois[[n]], layer="exprs", display="all")
  
  # Plot simulated vs true composite measurements  
  img <- plot_grid(plotlist=img, ncol=2, 
                   labels=unlist(lapply(X_simulation.list, function(sce){metadata(sce)$ground_truth})),
                   label_size=15, hjust=c(-2, -1.5), 
                   vjust=1, scale=0.9)
  print(img)
  cat("\n\n")
}

# Delete objects no longer needed (since there are a lot of big objects in this RMD)
rm(X_simulation.list)
```


## Training vs Experiment
***

### Scatterplot of Correlations per Protein *(arcsinh)*

Scatterplot of correlations between ground truth and simulated or decomposed results 
per protein for arcsinh transformed counts .

```{r results_training_vs_experiment, fig.asp=1, fig.width=4, fig.align="center"}
aoi <- "exprs"

# Calculate correlations between ground truth and simulated data for each protein
X.cor <- X.cor %>% 
  dplyr::rename(correlation_decomposition=correlation) %>%
  merge(lapply(keep(X_results.list, function(x){
    metadata(x)$comparison=="training"}), 
    function(sce){
      counts.long <- as.data.frame(assays(sce)[[aoi]]) %>%
        mutate(protein=rownames(.)) %>%
        melt() %>%
        dplyr::rename(!!as.symbol(metadata(sce)$ground_truth):=value,
                      cell=variable) 
      
    }) %>% bind_rows() %>%
      group_by(protein, cell) %>%
      summarise_all(na.omit) %>%
      group_by(protein) %>%
      summarise(correlation_training=cor(ground_truth, simulated)))


# Plot correlations for each protein
correlations.plot <- ggscatter(X.cor,
                               x="correlation_training", 
                               y="correlation_decomposition",
                               label="protein",
                               label.rectangle=TRUE,
                               color=pal_npg("nrc")("1"),
                               add.params=list(color=pal_npg("nrc")("4")[4],
                                               size=2)) +
  theme_cowplot(title.fontsize) +
  geom_abline(slope=1, linetype="dashed") +
  stat_cor(size=2) +
  ylim(0, 1) +
  xlim(0, 1)

print(correlations.plot)
```


## Decomposed vs Ground truth Measurements
***

### Images {.tabset}

Here we show the decomposed measurements vs the ground truth measurements for 
all images. 

```{r plot_images_decomposed, fig.asp=0.25, fig.width=18, fig.align="center", fig.keep=c(3, 6, 9, 12, 15, 18, 21, 24), results="asis"}
# Subset list to decomposed and ground truth expressions
X_decomposition.list <- keep(X_results.list, function(x){
  metadata(x)$comparison=="experiment"})
# Find the names of all protein channels
pois <- rownames(X_decomposition.list[[1]])
names(pois) <- rownames(X_decomposition.list[[1]])


# Call plot_cells to get individual plots for all rois for decomposed and real
# measurements
for (n in names(pois)){
  cat("####", n, "\n")
  
  img <- plot_cells(X_decomposition.list, masks, pois[[n]], layer="exprs", display="all")

  # Plot decomposed vs true results 
  img <- plot_grid(plotlist=img, ncol=2, 
                   labels=unlist(lapply(X_decomposition.list, function(sce){metadata(sce)$ground_truth})),
                   label_size=15, hjust=c(-2, -1.5), 
                   vjust=1, scale=0.9)
  print(img)
  cat("\n\n")
}
```

### Correlations per Protein *(arcsinh)*

Correlation between ground truth and decomposed results per protein for  
arcsinh transformed counts.

```{r results_per_protein_cor, fig.asp=1.2, fig.width=5, fig.align="center"}
aoi <- "exprs"
# Calculate correlations between ground truth and decomposed data for each protein
X.cor <- lapply(X_decomposition.list, function(sce){
  counts.long <- as.data.frame(assays(sce)[[aoi]]) %>%
    mutate(protein=rownames(.)) %>%
    melt() %>%
    dplyr::rename(!!as.symbol(metadata(sce)$ground_truth):=value,
                  cell=variable)
}) %>% bind_rows() %>%
  group_by(protein, cell) %>%
  summarise_all(na.omit) %>%
  group_by(protein) %>%
  summarise(correlation=cor(ground_truth, decomposed))

# Plot correlations per protein
proteinCorr <- plot_protein_cor(X.cor) + ylim(0, 1)
print(proteinCorr)
```

### Clustering

In the next part, we have a look at different clustering results. For this, we first 
cluster the ground truth and decomposed results together and show the results in a 
UMAP. We the repeat this for the argsinh and min-max transformed counts. Next, we 
look at the results if we use batch correction on the decomposed results. Lastly, 
we use Rhenogrpaph individually on the decomposed and ground truth data and compare 
the results.

#### UMAP of Combined Clustering of Ground Truth and Decomposed Data {.tabset}

```{r comb_clustering}
# Rename cells in experiment list before merging into single SCE object
X_decomposition_renamed.list <- lapply(X_decomposition.list, function(sce.temp){
  colnames(sce.temp) <- paste(colnames(sce.temp),
                              metadata(sce.temp)$ground_truth, sep="\n")
  colData(sce.temp)$ground_truth <- rep(metadata(sce.temp)$ground_truth, 
                                        nrow(colData(sce.temp)))
  sce.temp
})
# Add decomposed and ground truth data into one SCE
X.combined <- do.call("cbind", X_decomposition_renamed.list)

# Delete objects no longer needed (since there are a lot of big objects in this RMD)
rm(X_decomposition_renamed.list)

# Cluster combined SCE using rphenograph with default parameters
mat.combined <- t(assay(X.combined, "exprs"))
out.combined <- Rphenograph(mat.combined, k=30)
clusters.combined <- factor(membership(out.combined[[2]]))

# Add clustering to SCE
X.combined$clusters <- clusters

# Add reduced dimensions
set.seed(220225)
X.combined <- lapply(X.combined, function(sce.temp){
  # Add arcsinh and min-max transformed counts
  assay(sce.temp, "exprs") <- asinh(counts(sce.temp)/1)
  counts.t <- as.data.frame(t(counts(sce.temp)))
  assay(sce.temp, "min-max") <- t(predict(preProcess(counts.t, method=c("range")), 
                                          counts.t))
  
  # Run UMAP on raw counts, arsinh and min-max transformed counts
  sce <- runUMAP(sce, exprs_values="counts", name="UMAP_counts") 
  sce <- runUMAP(sce, exprs_values="exprs", name="UMAP_exprs") 
  sce <- runUMAP(sce, exprs_values="min-max", name="UMAP_min-max")
  sce
})
```

```{r plot_comb_clustering, fig.asp=0.8, fig.width=10, fig.align="center", results="asis"}
# Plot UMAP's
for (n in names(reducedDim(X.combined))){
  cat("#####", n, "\n")
  
  umap.combined <- dittoDimPlot(X.combined,
                                var="ground_truth",
                                reduction.use=n,
                                size=0.2,
                                color.panel=pal_npg("nrc")("2")) +
    theme_cowplot() +
    theme(legend.position="bottom") +
    guides(color=guide_legend(override.aes=list(size=axis_title.fontsize)))
  
  print(umap.combined)
  cat("\n\n")
}
```

#### Batch Correction of Decomposed Data using Ground Truth

```{r batch_correction}
set.seed(220228)
out.mnn <- fastMNN(X.combined, batch=X.combined$ground_truth,
               auto.merge=TRUE,
               assay.type="exprs",
               k=30)

# Transfer the correction results to the main spe object
reducedDim(X.combined, "fastMNN") <- reducedDim(out.mnn, "corrected")

# Get 30 nearest neighbours and count for how many cells in ground truth the 
# decomposed cell is in there
nn_30 <- reducedDim(out.mnn, "merge.info")$pairs

```

`r fds`/`r fdsaf` cells in the ground truth data have their corresponding decomposed 
cell in their 30 nearest neighbours when using fastMNN() for batch correction.

```{r plot_batch_correction, fig.asp=0.8, fig.width=10, fig.align="center"}
# Add reduced fastMNN corrected dimensions
set.seed(220225)
X.combined <- runUMAP(X.combined, dimred="fastMNN", name="UMAP_mnnCorrected")

# Plot fastMNN corrected UMAP
umap.combined <- dittoDimPlot(X.combined,
                              var="ground_truth",
                              reduction.use="UMAP_mnnCorrected",
                              size=0.2,
                              color.panel=pal_npg("nrc")("2")) +
  theme_cowplot() +
  theme(legend.position="bottom") +
  guides(color=guide_legend(override.aes=list(size=axis_title.fontsize)))

print(umap.combined)
```

#### Individual Clustering of Ground Truth and Decomposed Data {.tabset}

```{r individual_clustering, fig.asp=1, fig.width=15, fig.align="center", results="asis"}
# Cluster ground truth and decomposed SCE using rphenograph with default parameters 
# and add to SCE objects
X_decomposition.list <- lapply(X_decomposition.list, function(sce.temp){
  mat <- t(assay(sce.temp, "exprs"))
  out <- Rphenograph(mat, k=30)
  sce.temp$clusters <- factor(membership(out[[2]]))
  
  sce.temp
})
# Put all clusters for ground truth and decomposed data for each cell into one 
# data frame for easier handling later on
all.clusters <- lapply(X_decomposition.list, function(sce.temp){
  clusters <- sce.temp$clusters %>%
    as.data.frame() %>%
    dplyr::rename(!!as.symbol(metadata(sce.temp)$ground_truth):=".") %>%
    mutate(cell=colData(sce)$cell_id) 
  
  clusters
}) %>% bind_rows() %>%
  group_by(cell) %>%
  summarise_all(na.omit) %>%
  ungroup()

# Plot expression heatmaps for ground truth and decomposed data
for (sce in X_decomposition.list){
  cat("#####", metadata(sce)$ground_truth, "\n")
  
  cluster.expr <- dittoHeatmap(sce, 
                               assay="exprs", 
                               scale="none",
                               heatmap.colors=magma(100), 
                               annot.by=c("clusters"))
  
  print(cluster.expr)
  cat("\n\n")
}
```

```{r individual_clustering, fig.asp=1, fig.width=15, fig.align="center", results="asis"}
# Annotate clusters with celltypes
celltypes.gt <- data.frame(ground_truth=unique(all.clusters$ground_truth))
celltypes.gt$celltypes_gt <- c()
celltypes.dec <- data.frame(decomposed=unique(all.clusters$decomposed))
celltypes.dec$celltypes_decomposed <- c()

# Add celltypes to clusters dataframe
all.clusters <- all.clusters %>%
  merge(celltypes.gt) %>%
  merge(celltypes_decomposed)

# Add celltypes to SCE objects
X_decomposition.list <- lapply(X_decomposition.list, function(sce.temp){
  if (metadata(sce.temp)$ground_truth=="ground_truth"){
    colData(sce.temp) <- colData(sce.temp) %>%
      merge(celltypes.gt, by.x="clusters", by.y="ground_truth")
  } else {
    colData(sce.temp) <- colData(sce.temp) %>%
      merge(celltypes.dec, by.x="clusters", by.y="decomposed")
  }
})

# Summarize how many ground truth cells of a certain celltype have the same celltype 
# in the decomposed data (divided by the cluster size)
clusters.overlap <- table(all.clusters %>% 
                        pull(celltypes_decomposed),
                      all.clusters %>% 
                        pull(celltypes_gt)) %>%
  scale(center=FALSE, scale=colSums(.))

# Plot how many cells have a specified celltype in ground truth data and the same 
# celltype in the decomposed data divided by the cluster size of the ground truth data
cluster_overlap.plot <- Heatmap(clusters.overlap,
                           show_row_dend=FALSE,
                           show_column_dend=FALSE,
                           col=rev(magma(100)), # colorRamp2(c(0, 0.1, 5), rev(magma(3))), 
                           row_names_gp=gpar(fontsize=axis_title.fontsize),
                           column_names_gp=gpar(fontsize=axis_title.fontsize),
                           column_title_gp=gpar(fontsize=title.fontsize),
                           rect_gp=gpar(col="white", lwd=1)) 
plot(cluster_overlap.plot)
```


## Matrix designs
***

### U 

```{r plot_U, fig.asp=0.7, fig.width=8, fig.align="center"}
# Plot U 
u.plot <- plot_single_U(u, paste0("U\n(dictionary size: ", length(unique(u$module)), ")"))
print(u.plot)
```

### A 

```{r plot_A, fig.asp=0.6, fig.width=8, fig.align="center"}
# Plot A 
a.plot <- plot_single_A(a, "A")
print(a.plot)
```